An equivalence of compositional frameworks

### 2018-10-13 EquivCompFrameworks_20180911kc.tex
added a bit about the old decorated cospans version of open Markov processes

Let $\mathrm{Graph}_k$ be the category whose objects are given by $k$-graphs and morphisms by morphisms of $k$-graphs, where a morphism of $k$-graphs is given by a pair of functions $f \colon E \to E^\prime$ and $g \colon V \to V^\prime$ between the edge sets and vertex sets, respectively, of two $k$-graphs that respect the source and target functions of each. In a previous work introducing structured cospans, the first two authors showed that the category $\mathrm{Graph}_k$ has finite colimits. We can then obtain a double category of open passive linear circuits by defining a left adjoint $L \colon \mathrm{Set} \to \mathrm{Graph}_k$ that assigns to a set $V$ the discrete passive linear circuit $L(V)$ given by passive linear circuit with $V$ as its set of vertices and no edges. The resulting symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathrm{Graph}_k)$ has:

@@

\node (E) at (4,0) {$\mathbb{E}$};
\node (F) at (5,0) {$\mathbb{V}$};

@@

\node (E) at (3,0) {$U$};
\node (F) at (5,0) {$W$};
\node (G) at (4,0) {$V$};
\node (E') at (3,-1) {$U^\prime$};
\node (F') at (5,-1) {$W^\prime$};
\node (G') at (4,-1) {$V^\prime$};
\node (A) at (6,0) {$K \in F(V)$};
\node (B) at (6,-1) {$K^\prime \in F(V^\prime)$};
(F) edge node[above]{$o$} (G)

@@

(E) edge node[above]{$i$} (G)
(E') edge node[above]{$i^\prime$} (G')
(F') edge node[above]{$o^\prime$} (G');
together with a morphism of $k$-graphs $\iota \colon F(h)(K) \to K^\prime$ in $F(V^\prime)$.}

@@

@@

\node (E) at (3,0) {$L(X)$};
\node (F) at (5,0) {$L(Y)$};
\node (G) at (4,0) {$P$};
\node (E') at (3,-1) {$L(X^\prime)$};
\node (F') at (5,-1) {$L(Y^\prime)$};
\node (G') at (4,-1) {$P^\prime$};
(F) edge node[above]{$O$} (G)

@@

(E) edge node[above]{$I$} (G)
(E') edge node[above]{$I^\prime$} (G')
(F') edge node[above]{$O^\prime$} (G');

@@

\node (E) at (3,0) {$X$};
\node (F) at (5,0) {$Y$};
\node (G) at (4,0) {$Z$};
\node (E') at (3,-1) {$X^\prime$};
\node (F') at (5,-1) {$Y^\prime$};
\node (G') at (4,-1) {$Z^\prime$};
\node (A) at (6,0) {$P \in F(Z)$};
\node (B) at (6,-1) {$P^\prime \in F(Z^\prime)$};
(F) edge node[above]{$o$} (G)

@@

(E) edge node[above]{$i$} (G)
(E') edge node[above]{$i^\prime$} (G')
(F') edge node[above]{$o^\prime$} (G');

>>>>

Let $\mathrm{Graph}_k$ be the category whose objects are given by $k$-graphs and morphisms by morphisms of $k$-graphs, where a morphism of $k$-graphs is given by a pair of functions $f \colon E \to E^\prime$ and $g \colon V \to V^\prime$ between the edge sets and vertex sets, respectively, of two $k$-graphs that respect the source and target functions of each. In a previous work introducing structured cospans, the first two authors showed that the category $\mathrm{Graph}_k$ has finite colimits \cite{BC2}. We can then obtain a double category of open passive linear circuits by defining a left adjoint $L \colon \mathrm{Set} \to \mathrm{Graph}_k$ that assigns to a set $V$ the discrete passive linear circuit $L(V)$ given by passive linear circuit with $V$ as its set of vertices and no edges. The resulting symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathrm{Graph}_k)$ has:

@@

\node (E) at (4,0) {$E$};
\node (F) at (5,0) {$V$};

@@

\node (E) at (3,0) {$U_1$};
\node (F) at (5,0) {$W_1$};
\node (G) at (4,0) {$V_1$};
\node (E') at (3,-1) {$U_2$};
\node (F') at (5,-1) {$W_2$};
\node (G') at (4,-1) {$V_2$};
\node (A) at (6,0) {$K_1 \in F(V_1)$};
\node (B) at (6,-1) {$K_2 \in F(V_2)$};
(F) edge node[above]{$o_1$} (G)

@@

(E) edge node[above]{$i_1$} (G)
(E') edge node[above]{$i_2$} (G')
(F') edge node[above]{$o_2$} (G');
together with a morphism of $k$-graphs $\iota \colon F(h)(K_1) \to K_2$ in $F(V_2)$.}

@@

In another previous work \cite{BFP}, Fong, Pollard and the first author used decorated cospans to construct a symmetric monoidal category of `open Markov processes'. In this framework, a Markov process on a finite set $N$ is given by a diagram in $\mathrm{Set}$:
\node (D) at (3,0) {$(0,\infty)$};
\node (E) at (4,0) {$E$};
\node (F) at (5,0) {$N$};
(E) edge [bend right] node [below] {$t$} (F);
where $E$ and $N$ are finite sets of edges and nodes, respectively. This is really just a special case of the previous example of passive linear circuits with $k = \mathbb{R}$. An open Markov process is then of course a cospan of finite sets where the apex is equipped with a Markov process:
\node (B) at (1,0) {$N$};
\node (E) at (4,0) {$E$};
\node (F) at (5,0) {$N$};
Fong, Pollard and the first author then add extra structure to open Markov processes such as populations at each node to obtain a symmetric monoidal category $\mathrm{DetlBalMark}$ of open Markov processes in `detailed balance' and then construct a black box functor $\blacksquare \colon \mathrm{DetBalMark} \to \mathrm{LinRel}$ that describes the steady state behavior of an open Markov process in detailed balance. On the way to doing this, one of the categories they construct using Fong's decorated cospan machinery is a symmetric monoidal category $\mathrm{Mark}$ which has:
\item{objects as finite sets and}
\item{morphisms as isomorphism classes of open Markov processes, where composition is by pushout.}
This is done using a symmetric lax monoidal functor $F \colon \mathrm{FinSet} \to \mathrm{Set}$ which assigns to each finite set $N$ the (large) set of all Markov processes on $N$ as defined above. Viewing this functor $F$ as now a symmetric lax monoidal pseudofunctor $F \colon \mathrm{FinSet} \to \mathrm{Cat}$ that assigns to a finite set $N$ the \emph{category} $F(N)$ of all Markov processes on $N$, we then get by Theorem \ref{DC} a symmetric monoidal double category $\mathbb{F}\textnormal{Cospan}(\mathrm{FinSet})$ which has:
\item{open Markov processes as horizontal 1-cells, and}
\item{maps of open Markov processes as 2-morphisms which are given by maps of cospans of finite sets:
\node (B) at (1,0) {$N_1$};
\node (D) at (1,-1) {$N_2$};
\node (E) at (3,0) {$M_1 \in F(N_1)$};
\node (E') at (3,-1) {$M_2 \in F(N_2)$};
(C) edge node[left]{$f^\prime$} (C')
together with a map of Markov processes $\iota \colon F(h)(M_1) \to M_2$ in $F(N_2)$, where a map between two Markov processes to be given by a pair of functions $(g,h)$ that make the following diagram commute:
\node (E) at (3,0) {$(0,\infty)$};
\node (E') at (3,-1) {$(0,\infty)$};
\node (G) at (5,0) {$N_1$};
\node (G') at (5,-1) {$N_2$};
(F) edge node [left] {$g$} (F')
(G) edge node [right] {$h$} (G');

In a more recent work \cite{BC}, the first two authors have constructed a symmetric monoidal double category of `open Markov processes' and `coarse-grainings', where roughly speaking, a coarse-graining is a way of approximating a larger open Markov process by a smaller one by partitioning the set of states into `lumps'. This construction uses neither decorated cospans nor structured cospans.

@@

\node (E) at (3,0) {$L(X_1)$};
\node (F) at (5,0) {$L(Y_1)$};
\node (G) at (4,0) {$P_1$};
\node (E') at (3,-1) {$L(X_2)$};
\node (F') at (5,-1) {$L(Y_2)$};
\node (G') at (4,-1) {$P_2$};
(F) edge node[above]{$O_1$} (G)

@@

(E) edge node[above]{$I_1$} (G)
(E') edge node[above]{$I_2$} (G')
(F') edge node[above]{$O_2$} (G');

@@

\node (E) at (3,0) {$X_1$};
\node (F) at (5,0) {$Y_1$};
\node (G) at (4,0) {$Z_1$};
\node (E') at (3,-1) {$X_2$};
\node (F') at (5,-1) {$Y_2$};
\node (G') at (4,-1) {$Z_2$};
\node (A) at (6,0) {$P_1 \in F(Z_1)$};
\node (B) at (6,-1) {$P_2 \in F(Z_2)$};
(F) edge node[above]{$o_1$} (G)

@@

(E) edge node[above]{$i_1$} (G)
(E') edge node[above]{$i_2$} (G')
(F') edge node[above]{$o_2$} (G');
